\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-04-20 future.tex
updates

Partitioned inference and decision-making within \rx also lends itself
to graceful system degradation, making the overall approach effective
towards increased robustness. There are two outstanding diagnosis and
failure recovery challenges that we need to address; the first has to
do with methodology related to model design of the instrument payload
for a highly modular and configurable vehicle; this is local to a
reactor. Doing so would enable (as a first step) using systematic
Fault Diagnosis, Isolation and Recovery (FDIR) algorithms as
demonstrated in the \texttt{RAX} experiment
\cite{williams96,mus98,williams97} and more recently on an AUV by
Dearden and others \cite{wang09,ernits10,dearden11}. The second
challenge has to implementing a policy within the \rx framework that
explicitly deals with recovery since controller failure can be
distributed across reactors. Since reactors are hierarchical in their
relationship to one another, their well defined dependencies allow
them to be removed from an agent systematically. Reactors more
abstract in scope can be removed before those which are less abstract
(and closer to the hardware). In the event a Navigation reactor, for
example, has a fault, it can be removed safely with lower level
functionalities (such as those in the executive). When the Navigator
is removed, it will recursively remove more abstract reactors (such as
a Mission Manager) that depend on it. Such a safety feature then
allows the system designers to ensure that early on they can invest
more in ensuring that reactors lower in the hierarchy are robust and
potentially formally verified. It also means that there is a safe and
sure way to take control of the vehicle at any level within the
reactors hierarchy; as long as low level reactors are still active one
will always been able to send goals/commands to them and execute what
are traditionally known as runout sequences (or clean up scripts).

Feature recognition for Sampling is a general research thrust; however
it has specific relevance in the context of \rx and AUV control. As
noted in Section \ref{sec:results} we have already demonstrated
Machine Learning techniques for recognition of select features in the
coastal ocean. A principal challenge that remains is that of obtaining
expert labeled data for any supervised or semi-supervised learning
methods.  While supervised techniques for learning classifiers such as
decision trees (\cite{Quinlan93-dtrees}), instance-based
classification (\cite{Aha-ibl-ml91}), Bayesian classifiers
(\cite{Jensen2001-BNetworks}) and artificial neural networks
(\cite{ANNsurvey-2000}) are popular, given the vast amount of data
available and poor understanding of correlation between physical and
biogeochemical variables in the coastal ocean, our work on INLs
\cite{mcgann08d,ryan10} and reliance on a oceanographic expert has
shown that scaling to different problems remains problematic. This is
a prime motivation to move towards more semi-supervised methods such
as \cite{kumar11,sergio12}. With the advent of the \od however,
another direction our research is taking us is in \texttt{Recommender
  Systems} \cite{Adomavicius05} as noted in Section
\ref{sec:results}. We are working to design and deploy software
infrastructure that will learn from user tagged data. To do so, we
will use tagging techniques similar to what users of commercial image
sharing sites like Flickr and Picassa do, to learn the context of the
data (including remote sensing). In addition, the \od will integrate
statistical Machine Learning techniques with incoming asynchronous
data stream obtained by sensors and platforms. Identification models
will be based on existing training sets for INLs and plumes; models
for other features of scientific interests driven by \can requirements
will be built in addition. Existing AUV archives for instance provide
sufficient data for a learning base for a range of features of
interest to \can; in addition we will build software that will learn
to incorporate incoming data streams to provide an existence proof of
a feature of interest.

>>>>

Partitioned inference and decision-making within \rx also lends itself
to graceful system degradation, making the overall approach effective
towards increased robustness. There are two outstanding diagnosis and
failure recovery challenges that we need to address; the first has to
do with methodology related to model design of the instrument payload
for a highly modular and configurable vehicle; this is local to a
reactor. Doing so would enable (as a first step) using systematic
Fault Diagnosis, Isolation and Recovery (FDIR) algorithms as
demonstrated in the \texttt{RAX} experiment
\cite{williams96,mus98,williams97} and more recently on an AUV by
Dearden and others \cite{wang09,ernits10,dearden11}. The second
challenge has to \kcomment{do with} implementing a policy within the
\rx framework that explicitly deals with recovery since controller
failure can be distributed across reactors. Since reactors are
hierarchical in their relationship to one another, their well defined
dependencies allow them to be removed from an agent
systematically. Reactors more abstract in scope can be removed before
those which are less abstract (and closer to the hardware). In the
event a Navigation reactor, for example, has a fault, it can be
removed safely with lower level functionalities (such as those in the
executive). When the Navigator is removed, it will recursively remove
more abstract reactors (such as a Mission Manager) that depend on
it. Such a safety feature then allows the system designers to ensure
that early on they can invest more in ensuring that reactors lower in
the hierarchy are robust and potentially formally verified. It also
means that there is a safe and sure way to take control of the vehicle
at any level within the reactors hierarchy; as long as low level
reactors are still active one will always been able to send
goals/commands to them and execute what are traditionally known as
runout sequences (or clean up scripts).

Feature recognition for Sampling is a general research thrust; however
it has specific relevance in the context of \rx and AUV control. As
noted in Section \ref{sec:results} we have already demonstrated
Machine Learning techniques for recognition of select features in the
coastal ocean. A principal challenge that remains is that of obtaining
expert labeled data for any supervised or semi-supervised learning
methods.  While supervised techniques for learning classifiers such as
decision trees (\cite{Quinlan93-dtrees}), instance-based
classification (\cite{Aha-ibl-ml91}), Bayesian classifiers
(\cite{Jensen2001-BNetworks}) and artificial neural networks
(\cite{ANNsurvey-2000}) are popular, given the vast amount of data
available and poor understanding of correlation between physical and
biogeochemical variables in the coastal ocean, our work on INLs
\cite{mcgann08d,ryan10} and reliance on a oceanographic expert has
shown that scaling to different problems remains problematic. This is
a prime motivation to move towards more semi-supervised methods such
as \cite{kumar11}. With the advent of the \od however,
another direction our research is taking us is in \texttt{Recommender
  Systems} \cite{Adomavicius05} as noted in Section
\ref{sec:results}. We are working to design and deploy software
infrastructure that will learn from user tagged data. To do so, we
will use tagging techniques similar to what users of commercial image
sharing sites like Flickr and Picassa do, to learn the context of the
data (including remote sensing). In addition, the \od will integrate
statistical Machine Learning techniques with incoming asynchronous
data stream obtained by sensors and platforms. Identification models
will be based on existing training sets for INLs and plumes; models
for other features of scientific interests driven by \can requirements
will be built in addition. Existing AUV archives for instance provide
sufficient data for a learning base for a range of features of
interest to \can; in addition we will build software that will learn
to incorporate incoming data streams to provide an existence proof of
a feature of interest.

### 2012-04-20 results.tex
updates

A significant driver of the applicability of our work was the
growing need for AUVs mission to be driven by scientific sensor 
data.This trend is illustrated by the 
implementation of MBARI's \emph{Gulper} water-samplers \cite{Bird07},
$2$ litre syringe style containers $10$ of which are encased within
the mid-body section of our Dorado vehicle as shown in Fig
\ref{fig:gulper}. However, the original design and intended use of
operation continued to dominated by a naive belief of collecting and
returning water samples as traditional ship-borne methods often
employ, namely using pre-scripted methods, albeit on a highly capable
robot. With the installation of \rx onboard the Dorado with its
capability to synthesize plans and replan on the fly, more imaginative
methods of sample acquisition were targeted.

Our early work focused on observing and sampling within Intermediate
Nephaloid Layers (INLs) \footnote{INLs are fluid sheets of suspended
  particulate matter that originate from the sea floor
  \cite{mcphee-shaw2006}} \cite{ryan10}. The location and scale of
INLs in the water column, which depend upon seasonal and episodic
variability, are not predictable. In turn, they can be identified by
their high turbidity and low chlorophyll fluorescence. Initial
objectives focused on ways to enhance survey results by ensuring as
large a coverage as possible so that INL fields were observed at
tighter spaced transects and when no INL is detected, then transect
spacing in the lawn mower can be incrementally increased. Follow on
work directly targeted sample acquisition within INLs which are often
sparse, patchy in large horizontal scales (in Kms) and small vertical
scales (in meters). Fig. \ref{fig:inl} shows one example from an AUV
transect from 2005. The intent of both observation and sample
acquisition was to interrupt the nominal plan-execution cycle and
inject an action to alter and re-factor the plan dynamically; in one
instance altering the navigation, the other in actuating a sampler.

@@

While the early experiments in \cite{fox2007} coupled the classifier
with the onboard planner, two further novel extensions to sampling
were added. First, to enhance the classifier model by adding a
temporal dimension with a Hidden Markov Model (HMM) \cite{rabiner86,
  Rabiner89atutorial}.  An HMM is a sequential state transition model
in which the current state is not directly visible, but an
observation, dependent on the state, is.  Each state has a probability
distribution over the possible observations.  Therefore the sequence
of observations gives some information about the sequence of traversed
states in the HMM. Unlike identification models based on the simple
classification scheme, HMMs use the dynamics of transition between
modeled states to exploit sequential information of the sensor
readings providing more robust formulation for environmental state
identification. The temporal probabilistic transition then acts as a
filter on top of the clustering model since it is not only derived
from it, but it informs the planner of a potential trigger state only
when the models' transition occurs to an appropriate state of
recognition of the INL feature \cite{mcgann08b}. Not only was
deliberation aided by direct perception of the state (of the INL) but
temporal transition functions added a further level of validation
between expected states in the INL model. This coupling of Machine
Learned model with a deliberative planner is novel in the ocean
sciences and is a continuing direction of investigation \cite{kumar11,
  sergio12}.

@@

To augment our understanding of such microbial communities, the Dorado
with \rx has been utilized in a range of activities in a
quasi-lagrangian mode. As the water mass is being tracked with a
drifting instrument, the vehicle is expected to move around it. While
our model suppirt multiples patterns(square, lawn-mower,
transect(s)\dots) with any possible dimension factor, the
majority of at-sea deployments have used a square pattern ($3X3$
km\textsuperscript{2} or a $1X1$ km\textsuperscript{2})\footnote{Often
  post-hoc reconstruction of the data set for visualization drives how
  much adaptation the vehicle can be allowed to undertake.}. Plan
synthesis occurs frequently, but within the context of a drifting
point generated by a drifter. Fig. \ref{fig:dta-setup} shows the
general setup of our experiments which can be tailored with the use of
various computational methods within the \texttt{DTA} app which is
contained within the Oceanographic Decision Support System (\od; see
below) \cite{das11}. GPS data from the free-floating drifter is sent
via satellite and processed onshore to determine a projection (speed
and bearing) of the drifter. This data is sent to the AUV; at the
completion of the current survey, the vehicle receives this data as a
high-level goal which is decomposed and synthesized as a survey
plan. Currently, the vehicle completes the survey without any further
dynamic updates from the drifter. In order to minimize the range of
errors that the survey needs to encapsulate (data latency or surface
or sub-surface currents for those drifters which have a drogue) the
estimation is a projection in time and space of the drifters centroid
sent with this goal from the shore-side app.

>>>>

A significant driver of the applicability of our work was the growing
need for AUV missions to be driven by sensor data. This trend is
illustrated by the implementation of MBARI's \emph{Gulper}
water-samplers \cite{Bird07}, $2$ litre syringe style containers $10$
of which are encased within the mid-body section of our Dorado vehicle
as shown in Fig \ref{fig:gulper}. However, the original design and
intended use of operation continued to be dominated by a naive belief
of collecting and returning water samples as traditional ship-borne
methods often employ, namely using pre-scripted methods. With the
installation of \rx onboard the Dorado with its capability sense the
environment and to appropriately synthesize plans and replan on the
fly, more imaginative methods of sample acquisition were targeted.

Our early work focused on observing and sampling within Intermediate
Nephaloid Layers (INLs) \footnote{INLs are fluid sheets of suspended
  particulate matter that originate from the sea floor
  \cite{mcphee-shaw2006}.} \cite{ryan10}. The location and scale of
INLs in the water column, which depend upon seasonal and episodic
variability, are not predictable. In turn, they can be identified by
their high turbidity and low chlorophyll fluorescence. Initial
objectives focused on ways to enhance survey results by ensuring as
large a coverage as possible so that INL fields were observed at
tighter spaced transects and when no INL is detected, then transect
spacing in the lawn mower can be incrementally increased. Follow on
work directly targeted sample acquisition within INLs which are often
sparse, patchy in large horizontal scales (in Kms) and small vertical
scales (in meters). Fig. \ref{fig:inl} shows one example from an AUV
transect from 2005. The intent of both observation and sample
acquisition was to interrupt the nominal plan-execution cycle and
inject an action to alter and re-factor the plan dynamically; in one
instance altering the navigation, the other in actuating a sampler.

@@

While the early experiments in \cite{fox2007} coupled the classifier
with the onboard planner, two further novel extensions to sampling
were added. First, to enhance the classifier model by adding a
temporal dimension with a Hidden Markov Model (HMM) \cite{rabiner86,
  Rabiner89atutorial}.  An HMM is a sequential state transition model
in which the current state is not directly visible, but an
observation, dependent on the state, is.  Each state has a probability
distribution over the possible observations.  Therefore the sequence
of observations gives some information about the sequence of traversed
states in the HMM. Unlike identification models based on the simple
classification scheme, HMMs use the dynamics of transition between
modeled states to exploit sequential information of the sensor
readings providing more robust formulation for environmental state
identification. The temporal probabilistic transition then acts as a
filter on top of the clustering model since it is not only derived
from it, but it informs the planner of a potential trigger state only
when the models' transition occurs to an appropriate state of
recognition of the INL feature \cite{mcgann08b}. Not only was
deliberation aided by direct perception of the state (of the INL) but
temporal transition functions added a further level of validation
between expected states in the INL model. This coupling of Machine
Learned model with a deliberative planner is novel in the ocean
sciences and is a continuing direction of investigation
\cite{kumar11}.

@@

To augment our understanding of such microbial communities, the Dorado
with \rx has been utilized in a range of activities in a
quasi-lagrangian mode. As the water mass is being tracked with a
drifting instrument, the vehicle is expected to move around it. While
our model \kcomment{supports} multiples patterns (square, lawn-mower,
transect(s)\dots) with any possible dimension factor, the majority of
at-sea deployments have used a square pattern ($3X3$
km\textsuperscript{2} or a $1X1$ km\textsuperscript{2})\footnote{Often
  post-hoc reconstruction of the data set for visualization drives how
  much adaptation the vehicle can be allowed to undertake.}. Plan
synthesis occurs frequently, but within the context of a drifting
point generated by a drifter. Fig. \ref{fig:dta-setup} shows the
general setup of our experiments which can be tailored with the use of
various computational methods within the \texttt{DTA} app which is
contained within the Oceanographic Decision Support System (\od; see
below) \cite{das11}. GPS data from the free-floating drifter is sent
via satellite and processed onshore to determine a projection (speed
and bearing) of the drifter. This data is sent to the AUV; at the
completion of the current survey, the vehicle receives this data as a
high-level goal which is decomposed and synthesized as a survey
plan. Currently, the vehicle completes the survey without any further
dynamic updates from the drifter. In order to minimize the range of
errors that the survey needs to encapsulate (data latency or surface
or sub-surface currents for those drifters which have a drogue) the
estimation is a projection in time and space of the drifters centroid
sent with this goal from the shore-side app.

### 2012-04-20 setobook.tex
updates

\documentclass[graybox,12pt]{svmult}

@@

\abstract{We describe a general purpose Artificial Intelligence based
  control architecture that incorporates in-situ decision making for
  adaptation and control for autonomous underwater vehicles
  (AUVs). The Teleo-Reactive EXecutive (\texttt{T-REX}) framework
  deliberates about future states, plans for actions and executes
  generated activities while monitoring plans for anomalous
  conditions. Plans are no longer scripted a priori but synthesized
  onboard with high-level directives instead of low level
  commands. Further, the architecture uses multiple control loops for
  a 'divide and conquer' problem solving strategy allowing for
  incremental computational model building, robust and focused failure
  recovery, ease of software development and ability to use legacy or
  non-native computational paradigms. Vehicle adaptation and sampling
  occurs in situ with additional modules which can be selectively used
  depending on the application in focus. Abstraction in problem
  solving allows different applications to be programmed relatively
  easily, with little to no changes to the core search engine thereby
  making software engineering sustainable. The representational
  ability to deal with time and resources coupled with Machine
  Learning techniques for event detection allows balancing shorter
  term benefits with longer term needs, an important need as AUV
  hardware becomes more robust allowing persistent ocean sampling and
  observation. \rx is in routine operational use at MBARI, providing
  scientists a new tool to sample and observe the dynamic coastal
  ocean.}

@@

Rajan and Py are funded by a block grant from the David and Lucile
Packard Foundation to MBARI. They thank their collaborators John Ryan,
Julio Harvey, Jim Bellingham and Robert Vrijenhoek at MBARI. In
addition they thank colleagues Martha del Alto, David Korsmeyer and
Steven Zornetzer at NASA Ames who made \eu available as open
source. The authors are indebted to Nicola Muscettola for his ideas
and early efforts in Constraint-based Temporal Planning at NASA Ames
which forms the basis of this work.
The authors also want to thank everybdy who has contributed to the 
development of \eu and its online documentation over the years, specially Ari Jonsson, Jeremy Frank, Conor McGann, Paul Morris,
Michael Iatauro, Matthew Boyce and Tristan Smith.

>>>>

% \documentclass[a4paper, 12pt]{article}
\documentclass[graybox,14pt]{svmult}

@@

\abstract{We describe a general purpose Artificial Intelligence based
  control architecture that incorporates in-situ decision making for
  autonomous underwater vehicles (AUVs). The Teleo-Reactive EXecutive
  (\rx) framework deliberates about future states, plans for actions
  and executes generated activities while monitoring plans for
  anomalous conditions. Plans are no longer scripted a priori but
  synthesized onboard with high-level directives instead of low level
  commands. Further, the architecture uses multiple control loops for
  a 'divide and conquer' problem solving strategy allowing for
  incremental computational model building, robust and focused failure
  recovery, ease of software development and ability to use legacy or
  non-native computational paradigms. Vehicle adaptation and sampling
  occurs in situ with additional modules which can be selectively used
  depending on the application in focus. Abstraction in problem
  solving allows different applications to be programmed relatively
  easily, with little to no changes to the core search engine thereby
  making software engineering sustainable. The representational
  ability to deal with time and resources coupled with Machine
  Learning techniques for event detection allows balancing shorter
  term benefits with longer term needs, an important need as AUV
  hardware becomes more robust allowing persistent ocean sampling and
  observation. \rx is in routine operational use at MBARI, providing
  scientists a new tool to sample and observe the dynamic coastal
  ocean.}

@@

Rajan and Py are funded by a block grant from the David and Lucile
Packard Foundation to MBARI. They thank their collaborators John Ryan,
Julio Harvey, Jim Bellingham and Robert Vrijenhoek at MBARI. In
addition they thank colleagues Martha del Alto, David Korsmeyer and
Steven Zornetzer at NASA Ames who made \eu available as open
source. The authors thank other members of the \eu team over the
years, including Ari J\'onsson, Jeremy Frank, Conor McGann, Paul
Morris, Michael Iatauro, Matthew Boyce and Tristan Smith.  Minh Bin Do
(NASA Ames), William Yew Teck Tan (NUS, Singapore) and Rishi Graham
(MBARI) provided comments on early drafts; we are grateful for their
help in making this manuscript more accessible. Finally, the authors
are indebted to Nicola Muscettola for his ideas and early efforts in
Constraint-based Temporal Planning at NASA Ames which forms the basis
of much of this work.
